Title Goes Here

### 2023-02-07 inst/paper/_results.qmd
add co-author comments round 2

@@

```{r}
#| label: fig-auth-plot-batch2
#| fig-cap: "(A) Number of samples in which each compound was detected in the first and second batch. (B) Quantity (ng) of each compound extracted from each sample in batch 2. The plot displays the extracted quantity across the three washes and final calculus extraction (calc). Each coloured line represents a different calculus sample."
#| fig-width: 8
#| fig-height: 6

@@

compound_detect_plot <- uhplc_calculus_long %>%
  #remove_missing() %>%
  filter(quant > 0) %>% 
  ggplot(aes(x = compound, fill = batch)) +
    #geom_jitter(width = 0.2) +
    #geom_boxplot() +
    geom_bar(position = "dodge") +
    #facet_wrap(~ batch) +
    theme_bw() +
    scale_fill_viridis_d() +
    scale_x_discrete(labels = compound_names) +
    theme(
      axis.title.x = element_blank(),
      axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1),
      legend.position = "none",
      panel.grid.major.x = element_blank()
      ) +
    labs(y = "Count")

@@

```{r}
#| label: fig-detection-preservation
#| fig-cap: "(A) Violin plot with overlaid box plots depicting the distribution of extracted quantities of each compound from batch 2 separated by state of preservation of the skeleton. (B) Extracted quantity (ng) of compound plotted against weights of the calculus samples from batch 2."
#| fig-height: 6
#| fig-width: 7

preservation_plot <- uhplc_calculus_long %>%
  filter(
    !is.na(preservation),
    quant > 0,
    batch == "batch2",
    #!compound %in% c("cbd", "cbn", "cocaine", "thc", "thca-a", "thcva"), # remove compounds not detected
    id != "MB329" # unusually high nicotine and cotinine quantity in batch 1
    ) %>% 
  ggplot(aes(x = preservation, y = quant)) +
    geom_violin(aes(fill = preservation), alpha = 0.6) +
    #geom_jitter(width = 0.2) +
    geom_boxplot(width = 0.2) +
    facet_wrap(~ compound, scales = "free_y", labeller = labeller(compound = compound_names), dir = "v") +
    theme_bw() +
    theme(
      legend.position = "none",
      panel.grid.major.x = element_blank(),
      axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1)
      ) +
    labs(x = "Preservation", y = "Quantity (ng)") +
    scale_fill_viridis_d()

weight_quant_plot <- uhplc_calculus_long %>%
  filter(
    #!is.na(preservation),
    batch == "batch2",
    quant > 0,
    #!compound %in% c("cbd", "cbn", "cocaine", "thc", "thca-a", "thcva"), # remove compounds not detected
    id != "MB329" # unusually high nicotine and cotinine quantity in batch 1
    ) %>% 
  ggplot(aes(x = weight, y = quant)) +
    geom_point(col = viridisLite::viridis(1)) +
    facet_wrap(~ compound, scales = "free_y", labeller = labeller(compound = compound_names), dir = "v") +
    theme_bw() +
    theme(axis.title.y = element_blank()) +
    labs(y = "Quantity (ng)", x = "Weight (mg)")

@@

```{r}
#| label: tobacco-accuracy-setup
#tobacco <- uhplc_calculus_demography %>%
# should only be in males
tobacco <- uhplc_calculus_long %>%
  filter(
    compound %in% c("nicotine", "cotinine"),
    sex == "m" | sex == "pm"
    ) %>% 
  mutate(detection = if_else(quant == 0, 0, 1)) # 0 = not detected; NA = not included in batch 2.

@@

The presence of pipe notch(es) in an individual and concurrent detection of nicotine
and/or cotinine is used as a crude indicator of the accuracy of the method. When
combining the results of both batches, the method was able to detect some form
of tobacco in
`r nrow(filter(tobacco_accuracy, pipe_notch == "Y", detection > 0))`
of `r nrow(filter(tobacco_accuracy, pipe_notch == "Y"))`
individuals with a pipe notch
(`r scales::percent(nrow(filter(tobacco_accuracy, pipe_notch == "Y", detection > 0)) / nrow(filter(tobacco_accuracy, pipe_notch == "Y")), accuracy = 0.1)`).
When also considering correct the absence of a tobacco alkaloid together with the absence
of a pipe notch, the accuracy of the method is
`r scales::percent(mean(tobacco_accuracy$correct, na.rm = T), accuracy = 0.1)`.
Accuracy in the old adult age category is
`r scales::percent(filter(accuracy_age, age == "old")$mean, accuracy = 0.1)`.

In the replicated samples only, tobacco detection was successful in
`r nrow(filter(replicated_accuracy, pipe_notch == "Y", detection > 0))`
out of
`r nrow(filter(replicated_accuracy, pipe_notch == "Y"))`
pipe smokers
(`r scales::percent(nrow(filter(replicated_accuracy, pipe_notch == "Y", detection > 0)) / nrow(filter(replicated_accuracy, pipe_notch == "Y")), accuracy = 0.1)`)
Including individuals with absence of a pipe notch and concurrent absence of
compounds as a correct identification, gives an overall accuracy of
`r scales::percent(mean(replicated_accuracy$correct, na.rm = T), accuracy = 0.1)`.

@@

The strongest point-biserial (Pearson) correlation correlations were a near-perfect
positive correlation between `r cor_statement(corr_tib_out, strength = "strong")`,
and moderate correlations between
`r cor_statement(corr_tib_out, strength = "moderate")` (@tbl-pearson).

@@

```{r}
#| label: fig-polycorr
#| fig-cap: "Plot of the polychoric correlations (*rho*). Larger circles and increased opacity indicates a stronger correlation coefficient."
polycorr$rho %>%
  ggcorrplot::ggcorrplot(
    method = "circle", type = "lower", 
    colors = c(viridisLite::plasma(1, begin = 0), "white", viridisLite::plasma(1, begin = 1)),
    legend.title = expression(rho)
    ) +
  scale_x_discrete(labels = prettified_names) +
  scale_y_discrete(position = "right", labels = prettified_names) +
  theme(
    legend.position = "left",
    legend.title = element_text(face = "italic")
  )
```

>>>>

The pattern we expect to see in authentic compounds representing compounds
trapped within the dental calculus, 
is a reduction in the quantity from wash 1 to wash 3 as potential surface contaminants
are washed off, and then a
spike in the final extraction when entrapped compounds are released and detected.

@@

```{r}
#| label: fig-auth-plot-batch2
#| fig-cap: "(A) Number of samples in which each compound was detected in the first and second batch. (B) Quantity (ng) of each compound extracted from each sample in batch 2. The plot displays the extracted quantity across the three washes and final calculus extraction (calc). Each coloured line represents a different calculus sample. CBD = cannabidiol; CBN = cannabinol; THC = tetrahydrocannabinol; THCA-A = tetrahydrocannabinolic acid A; THCVA = tetrahydrocannabivarin acid."
#| fig-width: 8
#| fig-height: 6

@@

compound_detect_plot <- uhplc_calculus_long %>%
  #remove_missing() %>%
  filter(quant > 0) %>% 
  ggplot(aes(x = compound, fill = batch)) +
    #geom_jitter(width = 0.2) +
    #geom_boxplot() +
    geom_bar(position = "dodge") +
    #facet_wrap(~ batch) +
    theme_bw() +
    scale_fill_viridis_d() +
    scale_x_discrete(labels = compound_names) +
    theme(
      axis.title.x = element_blank(),
      axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1),
      legend.position = "left",
      panel.grid.major.x = element_blank()
      ) +
    labs(y = "Count")

@@

```{r}
#| label: fig-detection-preservation
#| fig-cap: "(A) Violin plot with overlaid box plots depicting the distribution of extracted quantities of each compound from batch 2 separated by state of preservation of the skeleton. (B) Extracted quantity (ng) of compound plotted against weights of the calculus samples from batch 2."

preservation_plot <- uhplc_calculus_long %>%
  filter(
    !is.na(preservation),
    quant > 0,
    batch == "batch2",
    #!compound %in% c("cbd", "cbn", "cocaine", "thc", "thca-a", "thcva"), # remove compounds not detected
    id != "MB329" # unusually high nicotine and cotinine quantity in batch 1
    ) %>% 
  ggplot(aes(x = preservation, y = quant)) +
    geom_violin(aes(fill = preservation), alpha = 0.6) +
    #geom_jitter(width = 0.2) +
    geom_boxplot(width = 0.2) +
    facet_wrap(~ compound, scales = "free_y", labeller = labeller(compound = compound_names), dir = "v", ncol = 1) +
    theme_bw() +
    theme(
      legend.position = "none",
      panel.grid.major.x = element_blank(),
      axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1)
      ) +
    labs(x = "Preservation", y = "Quantity (ng)") +
    scale_fill_viridis_d()

weight_quant_plot <- uhplc_calculus_long %>%
  filter(
    #!is.na(preservation),
    batch == "batch2",
    quant > 0,
    #!compound %in% c("cbd", "cbn", "cocaine", "thc", "thca-a", "thcva"), # remove compounds not detected
    id != "MB329" # unusually high nicotine and cotinine quantity in batch 1
    ) %>% 
  ggplot(aes(x = weight, y = quant)) +
    geom_point(col = viridisLite::viridis(1)) +
    facet_wrap(~ compound, scales = "free_y", labeller = labeller(compound = compound_names), dir = "v", ncol = 1) +
    theme_bw() +
    theme(axis.title.y = element_blank()) +
    labs(y = "Quantity (ng)", x = "Weight (mg)")

@@

```{r}
#| label: tobacco-accuracy-setup
#tobacco <- uhplc_calculus_demography %>%
# should only be in males
tobacco <- uhplc_calculus_long %>%
  filter(
    batch == "batch2",
    compound %in% c("nicotine", "cotinine"),
    sex == "m" | sex == "pm"
    ) %>% 
  mutate(detection = if_else(quant == 0, 0, 1)) # 0 = not detected; NA = not included in batch 2.

@@

The presence of pipe notch(es) in an individual and concurrent detection of nicotine
and/or cotinine is used as a crude indicator of the accuracy of the method. Only
males were used in accuracy calculations, as pipe notches are ubiquitous in males,
but not in females.
In batch 2, the method was able to detect some form
of tobacco in
`r nrow(filter(tobacco_accuracy, pipe_notch == "Y", detection > 0))`
of `r nrow(filter(tobacco_accuracy, pipe_notch == "Y"))`
individuals with a pipe notch
(`r scales::percent(nrow(filter(tobacco_accuracy, pipe_notch == "Y", detection > 0)) / nrow(filter(tobacco_accuracy, pipe_notch == "Y")), accuracy = 0.1)`).
When also considering correct the absence of a tobacco alkaloid together with the absence
of a pipe notch, the accuracy of the method is
`r scales::percent(mean(tobacco_accuracy$correct, na.rm = T), accuracy = 0.1)`.
Accuracy in the old adult age category is
`r scales::percent(filter(accuracy_age, age == "old")$mean, accuracy = 0.1)`,
but with only `r sum(tobacco_accuracy$age == "old")` individuals.

@@

Point-biserial correlation was conducted on paired continuous and dichotomous
variables, to see if any relationships exist between extracted concentrations
and other variables.
The strongest point-biserial (Pearson) correlation correlations were a near-perfect
positive correlation between `r cor_statement(corr_tib_out, strength = "strong")`,
and moderate correlations between
`r cor_statement(corr_tib_out, strength = "moderate")` (@tbl-pearson).

@@

```{r}
#| label: fig-polycorr
#| fig-cap: "Plot of the polychoric correlations (*rho*). Larger circles and increased opacity indicates a stronger correlation coefficient. OA = osteoarthritis; VOP = vertebral osteophytosis; SN = Schmorl’s nodes; DDD = degenerative disc disease; CO = cribra orbitalia; CMS = chronic maxillary sinusitis; SA = salicylic acid."
polycorr$rho %>%
  ggcorrplot::ggcorrplot(
    method = "circle", type = "lower", 
    colors = c(viridisLite::plasma(1, begin = 0), "white", viridisLite::plasma(1, begin = 1)),
    legend.title = expression(rho)
    ) +
  scale_x_discrete(labels = prettified_names) +
  scale_y_discrete(position = "right", labels = prettified_names) +
  theme(
    legend.position = "left",
    legend.title = element_text(face = "italic")
  )
```
